Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increases Repair Time?

### 2015-02-26 phaseDelay.tex
kinda v0.1 semi-coherent

@@

@@

\begin{figure}
\begin{center}
\begin{tabular}{rrl}
year& \# issues&\\\hline
2006 &  44 &\\
2007 &  34 &\\
2008&  288 &\rule{1mm}{2mm}\\
2009&  846 &\rule{3mm}{2mm}\\
2010& 1007 &\rule{3mm}{2mm}\\
2011& 3273 &\rule{10mm}{2mm}\\
2012&18102 &\rule{45mm}{2mm}\\
2013&20336 &\rule{50mm}{2mm}\\
2014& 3307 & \rule{10mm}{2mm}\\\cline{1-2}
Total:&47228
\end{tabular}
\end{center}
\caption{This paper studies 47,228 issues recorded 2006 to 2014.}\label{fig:years}
\end{figure}
\begin{figure*}[!t]
\begin{center}
\includegraphics[width=6in]{waterfall2.png}
\end{center}
\caption{Different approaches to software development:  waterfall and agile (bottom left).}
\end{figure*}

@@

In these results, we checked if a small number of bugs are most expensive. 
To check this:
\bi
\item
We repeated the analysis that generated \fig{scale},
but instead of looking at the 50th percentile, we displayed the scale up factors
\item 
We only
checked the Design and Code scale up results since, from \fig{scale}, it is clear these
have the most examples of longest phase delays.
\ei 
The  results are shown in \fig{scale90} and these
results are somewhat different for Design and Coding issues:
\bi 
\item For Design issues,  these have the same
general form as the 50th percentile results. That is, while it it certainly faster
to remove thing sin the phase where they are created, once we leave that phase
it does not seem to matter much how many phases we wait before fixing the issue.
\item
\item For Coding issues, the seems little impact of phase delay on the time
required to fix.
\ei

@@

The data used in this work shirai14

The Team Software ProcessSM (TSPSM) provides a
framework that teams use to collect software process data in real
time, using a defined disciplined process.
In particular, three common features of the projects in our SEI TSP data set  are coaches,
peer review, and  personnel reviews:
\bi 
\item
A ``coach'' is the team member  authorized to submit project data
(before submission,
these coaches check the data for obvious errors);
\item 
``Personnel review'' is a technique taken  from the Personal Software
ProcessSM (PSPSM);
\item 
``Peer review'' is a standard technique in
traditional software engineering.
\ei
PSPSM encourages developers to continually make and review their personnel estimates
about their day-to-day tasks, then compare those estimates against the actual development effort.
In this way, developers can acquire a more realistic understanding of their work behaviour.
As to peer review,  Basili and Boehm write  commented in 2001~\cite{boehm01} 
that peer reviews can catch over half the defects introduced into a system.
Peer review can be conducted on any artifact generated anywhere in the software
lifecycle and can quickly be adapted to new kinds of artifacts.

@@

The phased used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases have the same  sub-activities of {\em review} and {\em inspect}\footnote{\bill{plz distinguish review and inspect. does one usually happen first?}}
\item Testing is divided into several statges \footnote{\bill{need a one
line description of test vs qualtest vs inttest vs systemTest vs AcceptTest. I tries some words in the Key to \fig{sedata}. dont know if i got it right.}}
\ei 
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

TSP is not antithetical to agile--  indeed a TSP-waterfall style project can adopt aspects
of agile.  For example the agile loop  could be applied
over one or more of the phases shown in the long waterfall chain of \fig{waterfall}. Some
TSP teams adopt something like test-driven-development\footnote{TDD is an agile-method
where the ``test'' is the primary driver of the design. The tests are written first,
then the code to support those tests. TDD proceeds in three steps: red (where there
are broken tests); green (where tests are passing); and, possibly, refactor (where
the code is re-organized based on feedback from those tests~\cite{fraser03}.} where
reviews are scheduled after testing. Some of the groups in the SEI TSP data used that approach but
they do not effect the main conclusions of this paper:
\bi 
\item They were strongly in the minority\bill{any numbers on this?};
\item Most of our phase delay data comes from much early in the waterfall model
of \fig{waterfall}.
\ei 
That said, there are some very ``un-agile'' aspects
of the processes used by the   projects in the TSP SEI data.
A TSP participant spends much time reflecting on the project,
undisturbed by the behaviour of the  executables.
In fact,
it can be days/weeks
before TSP participants gain  feedback from executing code, for the following reasons:  
\bi 
\item
For TSP projects, the times spent in the of a design activity/phase is   
approximately as much effort as the coding phase.  
\item Our TSP teams did not
 combine development and  testing. 
 \item  The TSP projects studied here offer new releases at least every three months, but
often much longer that that\bill{any numbers on that}. Note that this is a far slower
release schedule than seen in, say, continuous integration projects.  
\item The time devoted to personnel review'' and peer review (defined above)
is  about 50\% as much effort as the previous construction phase activity.

\ei

Having presented all that, we can now present the main point of this section:
\bi 
\item The results section of this paper fails to find that  phase delay causing dramatic increases
in time to fix issues;
\item This lack-of-phase-delay effect {\em cannot} be explained away just by  saying that
the projects in this sample adopted something like  agile methods to reduce re-work costs.
\ei

The Software Engineering Institute (SEI) at
Carnegie Mellon University has collected data from organizations
that have adopted TSP. 
Our efforts focused on data collected by projects that launched
after 2009 and used an automatic data recording tool. 
Due to confidentiality restrictions, we cannot offer too many details
on those project. That said, there were less than a dozen United States
Department of Defence projects. Also, the majority of
the projects were weither web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany) and a few from a commercial 
computer-aided design system (developed by a world wide distributed team).

Data collected included time logs, defect
logs, added and modified size logs, and work
breakdown strcture logs.
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).

In these logs
the log records include work start time, work end time, delta
work time, and interruption time. Software engineers are often
interrupted by meetings, requests for technical help, reporting, and
so forth. These events are recorded, in minutes, as interruption
time. In this paper, when we report ``time to resolve an
issue'', we show the difference between the start and end times
of a work session, with any interruption time subtracted (the
difference in times, minus the interruptions).

As of January 2014, the SEI TSP database contained data from 109
TSP projects. The projects started between July 2009 and
September 2013; they included 34 teams and 309 people. Among
the database fact tables, the time store contains 103,023 time logs,
18,408 defect logs, and 7,464 size logs\footnote{\bill{we need to 
update this table}}.

A common property of real-world data sets is the presence
of noisy entries (superfluous  or spurious data). 
The level of noise can be quite high. As reported
in \cite{shepperd12}, around
10\% to 30\%
of the records in the NASA MDP defect data sets are
affected by noise. Nichols et al.~\cite{shirai14}  report that
the noise levesl in the SEI TSE data are smaller than those seen
in other data sets. They found in the SEI TSP data that:\bi 
\item
4\% of the data was incorrect such as  null values of illegal formats;
\item  2\% of the data has inconsistencies such as timestamps
where the stop time was before the start time;
\item 3\% of the data had data that was not credible
such as tasks listed in one day that took like than six hours.
\ei 
That said, certain semantic features of the SEI TSP data should be noted.
Firstly, in the current TSP collection tool, 
fix times are only the developer time for the developer walking through the phases of \fig{waterfall}.
We are currently tracking the fix time for post-release issues (e.g. those raised during  acceptance test and later
product life cycle). So far, in that post-release data,  we have not detected
a dramatic phase escalation effect (but at this time, we have nothing definitive comment on that matter).

\bill{somewhere you have one note on \underline{find} and fix times.  for this paper, we need just fix times. but is there
anything we need to fret about re \underline{find} times?}

The majority of the teams follow a more linear process within a single feature of physical component (usually no more than a week or two work in total). Unlike some agile teams, *most*TSP teams will not conflate development and unit test.

The starting point for this paper were the times required to fix 47,228 issues.
All data was collected
by the Software Engineering Institute in the period 2006 to 2014 (see \fig{years}).
To the best of our knowledge, this is the largest sample of ``time to fix issues''
yet analyzed.

This data was collected using the   methodology described by
Watts Humphrey's Team Software Process~\cite{tsp00}. TSP is an extension of Humphrey's early
work on Personnel Software Process~\cite{psp05}.  PSP encourages developers to continually make estimates
about their day-to-day tasks, then compare those estimates against the actual development effort
(in this way, developers can acquire a more realistic understanding of their work behaviour).

In all data, was collected from 172 separate projects. Given Beck's comments (above) about phase delay motivating agile processes, we take care to identify the differences of these 
projects to agile development projects. An alternate, more traditional approach is a ``macro linear''
approach such as the ``waterfall methodology'' that develops code after an extensive
pre-planning stage:

The defect typ that we studied were
\be 
\item Environment: design, compile, test, or other support system problems;
\item Interface: procedure calls and reference, I/O, user format;
\item Data: structure, content; 
\item Documentation: comments, messages;
\item Syntax: spelling, punctuation typos, instruction formats
\item Function: logic, pointers, loops, recursion, computation, function defects  
\item Checking: error messages, inadequate checks;
\item Build: change management, library, version control;
\item Assignment: package
declaration, duplicate names, scope, limits;
\item System: configuration, timing, memory.
\ee

@@

>>>>

\begin{figure*}[!t]
\begin{center}
\includegraphics[width=6in]{waterfall2.png}
\end{center}
\caption{Different approaches to software development:  waterfall and agile (bottom left).}
\end{figure*}

@@

\input{data}

@@

@@

\section{Phase Delay in Slower Bugs?}
In these results, we checked if a small number of bugs are most expensive. 
To check this:
\bi
\item
We repeated the analysis that generated \fig{scale},
but instead of looking at the 50th percentile, we displayed the scale up factors
\item 
We only
checked the Design and Code scale up results since, from \fig{scale}, it is clear these
have the most examples of longest phase delays.
\ei 
The  results are shown in \fig{scale90} and these
results are somewhat different for Design and Coding issues:
\bi 
\item For Design issues,  these have the same
general form as the 50th percentile results. That is, while it it certainly faster
to remove thing sin the phase where they are created, once we leave that phase
it does not seem to matter much how many phases we wait before fixing the issue.
\item
\item For Coding issues, the seems little impact of phase delay on the time
required to fix.
\ei

@@

\section{Validity} 
A previous paper~\cite{shirai14} has applied a range of sanity checks to the data.
A common property of real-world data sets is the presence
of noisy entries (superfluous  or spurious data). 
The level of noise can be quite high. As reported
in \cite{shepperd12}, around
10\% to 30\%
of the records in the NASA MDP defect data sets are
affected by noise. Nichols et al.~\cite{shirai14}  report that
the noise levesl in the SEI TSE data are smaller than those seen
in other data sets. They found in the SEI TSP data that:\bi 
\item
4\% of the data was incorrect such as  null values of illegal formats;
\item  2\% of the data has inconsistencies such as timestamps
where the stop time was before the start time;
\item 3\% of the data had data that was not credible
such as tasks listed in one day that took like than six hours.
\ei

@@

@@

\begin{figure}
\begin{center}
\begin{tabular}{rrl}
year& \# issues&\\\hline
2006 &  44 &\\
2007 &  34 &\\
2008&  288 &\rule{1mm}{2mm}\\
2009&  846 &\rule{3mm}{2mm}\\
2010& 1007 &\rule{3mm}{2mm}\\
2011& 3273 &\rule{10mm}{2mm}\\
2012&18102 &\rule{45mm}{2mm}\\
2013&20336 &\rule{50mm}{2mm}\\
2014& 3307 & \rule{10mm}{2mm}\\\cline{1-2}
Total:&47228
\end{tabular}
\end{center}
\caption{This paper studies 47,228 issues recorded 2006 to 2014.}\label{fig:years}
\end{figure}
